Spatial distribution of environmental DNA in a nearshore marine habitat

### 2017-01-05 Manuscript/v1.1_PeerJ/response/response_to_reviews.tex

I think that it would be informative to add a paragraph in the discussion section commenting the caveats of using 16S as metabarcoding marker for taxonomic assignment of the MOTUs (i.e. the issues with the assignment of many sequences which can be assigned only to the family or lower levels), compared to other more variable markers with denser reference databases, such as COI, which, in most cases, allow assignments to the genus or species level.

\begin{response}
	We agree with your point that a locus with only coarse-scale resolution (16S) may obfuscate patterns of diversity compared to a locus with fine-scale resolution (CO1). A subset of us has recently published an in-depth comparison of multiple loci for metabarcoding, and TODO: we have added a more substantial comment on this and referenced to this now-published work.
\\
\end{response}

\subsubsection*{Specific Corrections}
Some specific corrections are detailed below:

L 36 to 40. This whole paragraph may be deleted to shorten the introduction, without any significant information loss.
\begin{response}
  We can understand the reviewer's perspective, but feel this point is a stylistic one and would prefer to leave the manuscript as it is currently worded to increase the appeal of our work to a broader audience.\\
\end{response}

@@

L90. Correct “Actiniaria” and “Bryozoans”.
\begin{response}
  Thanks for catching these errors. We made these changes.\\
\end{response}

L91. Delete “(Decapoda)”. These habitats hold many other Crustacea other than Decapoda as well. Actually, I would write “barnacles (Sessilia) and other crustaceans” in line 89.
\begin{response}
  Excellent suggestion; we made this change.\\
\end{response}

L93: “which provide”
\begin{response}
  We made this change.\\
\end{response}

L94: The microscopic plankton is not composed just by “diatoms and larvae”. Many important holoplanktonic organisms (e.g. copepods) could be mentioned here.
\begin{response}
  We did not state that the microscopic plankton is composed solely of diatoms and larve, we made a  statement and qualified it with two examples to give meaning and context: `diverse assemblage of microscopic plankton including diatoms and larvae'. We mention diatoms because they are a widely-known, numerically abundant, and ecologically important group that falls outside the scope of this study on metazoans. We mention larvae because it subtly hints at our major finding that our samples captured many organisms best known for their sessile, benthic stages (e.g. barnacles). Yes, we could add copepods to this list if you would like, but what qualifies a group for inclusion? Numerical abundance? Biomass?\\
\end{response}

@@

I felt that under the remit of the Ecological Analyses headings throughout the study, there could have been some focus on reporting the alpha diversity and rarefaction analysis in relation to whether effective sequencing depth had been achieved for the samples in question.
\begin{response}
  \\
\end{response}

\subsubsection{Comments for the Author}
I thought the paper was very interesting. The study was well designed, implemented, and written. It was useful to clarify the expectations and hypotheses outlined within, however, I felt more effort could have been made to provide recommendations on how some of the limitations in eDNA sampling could be overcome.
\begin{response}
  \\
\end{response}

The bioinformatics/statistical methods seemed to be very well thought out, and consistent with current best practices in the analyses of this type of sequencing data.
\begin{response}
  \\
\end{response}

%------------------------------------REVIEWER3----------------------------------
\subsection*{Comment on the preprint (Bernd Hänfling)}
This is well designed study and we would like to congratulate the authors to this data set and its excellent analysis. However, we feel that the authors have made an important omission in the discussion.

Given that the main focus of this paper is to explore the spatial distribution of community eDNA we feel that the authors should consider the results of our study on lake fish communities (Hänfling et al. 2016). In our paper we have used extensive spatial sampling of three transects across our main study lake with a total of 66 individual samples and to our knowledge there is no other data set on macrobial eDNA metabarcoding with a comparable extent of explicit sampling. Our paper confirms that the spatial distribution of individual species reflects their abundance but also shows that the pattern of distribution is not random and ecologically meaningful. Although there might be important differences between freshwater and marine habitats we feel that the results from our study are extremely relevant here and should not be omitted especially when claiming novelty of discussion point 2.

Bernd Hänfling, Lori Lawson Handley, Dan Read and Ian Winfield

Hänfling B, Lawson Handley L, Read DS, Hahn C, Li J, Nichols P, Blackman RC, Oliver A, Winfield IJ (2016) Environmental DNA metabarcoding of lake fish communities reflects long-term data from established survey methods. Molecular Ecology 25, 3101-3119.
\begin{response}
  \\
\end{response}

### 2017-01-05 Manuscript/v1.1_PeerJ/carkeek_grid.tex
correct affiliations to NMFS standards

% using package 'authblk':
\author[1]{James L. O'Donnell\thanks{jodonnellbio@gmail.com}}
\author[1]{Ryan P. Kelly} % rpkelly@uw.edu
\author[2]{Andrew O. Shelton} % ole.shelton@noaa.gov
\author[2]{Jameal F. Samhouri} % jameal.samhouri@noaa.gov
\author[1,3]{Natalie C. Lowell} % nclowell@uw.edu
\author[4]{Gregory D. Williams} % greg.williams@noaa.gov

% Make sure authors specify department, institution, and address.
\affil[1]{School of Marine and Environmental Affairs, University of Washington, 3707 Brooklyn Ave NE, Seattle, Washington 98105, USA}
\affil[2]{Northwest Fisheries Science Center, NOAA Fisheries, 2725 Montlake Blvd E, Seattle, Washington 98112, USA}
\affil[3]{School of Aquatic and Fishery Sciences, University of Washington, 1122 NE Boat St, Seattle, Washington 98105, USA}
\affil[4]{Pacific States Marine Fisheries Commission, Under contract to the Northwest Fisheries Science Center, NOAA Fisheries, 2725 Montlake Blvd E, Seattle, WA 98112}

>>>>

% using package 'authblk':
\author[1]{James L. O'Donnell\thanks{jodonnellbio@gmail.com}}
\author[1]{Ryan P. Kelly} % rpkelly@uw.edu
\author[2]{Andrew O. Shelton} % ole.shelton@noaa.gov
\author[3]{Jameal F. Samhouri} % jameal.samhouri@noaa.gov
\author[1,4]{Natalie C. Lowell} % nclowell@uw.edu
\author[5]{Gregory D. Williams} % greg.williams@noaa.gov

% Make sure authors specify department, institution, and address.
\affil[1]{School of Marine and Environmental Affairs, University of Washington, 3707 Brooklyn Ave NE, Seattle, Washington 98105, USA}
\affil[2]{Earth Resource Technology, Inc., Under contract to the Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA}
\affil[3]{Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd E, Seattle, Washington 98112, USA}
\affil[4]{School of Aquatic and Fishery Sciences, University of Washington, 1122 NE Boat St, Seattle, Washington 98105, USA}
\affil[5]{Pacific States Marine Fisheries Commission, Under contract to the Northwest Fisheries Science Center, National Marine Fisheries Service, National Oceanic and Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA}
